---
title: "Petr Mockup"
author: "Brian S. Yandell"
date: "3/17/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The motivation is to reduce memory usage. Matt told us that memory is his greatest obstacle, when running QTL browsers because he cannot add more `R/plumber` workers. Also, if you have a cheap Windows laptop you cannot even open many DO `.RData` projects simply because haplotype probability object is too big.
 
The solution would be to store data in some database and load just a slice you need at a time. I experimented with `SQLite` but it was too slow. Then I try feather and it run fine (R script attached – crunching 1000 SNPs at a time). It just does not use `qtl2scan` effectively because the package is not prepared for that.

Task: refactor this Rmd to use DOex and run some examples.

```{r}
library(dplyr)
library(qtl2convert)
library(qtl2geno)
library(qtl2scan)
library(qtl2plot)

library(feather)
```

```{r}
load("C:/Temp/kidney.RData")
```

Which gene is Ace?

```
idx <- which(annot.mrna$symbol == "Ace")
addcovar <- covar[,-1]

db.file <- "C:/Temp/genoprobs_kidney.feather"
```

```{r}
Nbatches = 1000
Nsnps <- nrow(snps)
# make SNPs split into Nbatches
batches <-  split(1:Nsnps, ceiling(seq_along(1:Nsnps)/Nbatches))
```

LOD object to be produced

```{r}
result <- NULL

for (i in 1:length(batches)) {

  # fetch data from the database
  tmp <- read_feather(db.file, batches[[i]])
  print(i)
  
  # dimnames harcoded, ADD ASSERTIONS LATER 
  tmp.snps <- names(tmp)
  tmp.probs <- as.array(as.matrix(tmp))
  dim(tmp.probs) <- c(nrow(tmp.probs) %/% 8, 8, ncol(tmp.probs))
  dimnames(tmp.probs) <- list(annot.samples$Mouse.ID, LETTERS[1:8], tmp.snps)
  
  # convert to qtl2 format
  probs <- probs_doqtl_to_qtl2(tmp.probs, snps[tmp.snps,], pos_column = "bp")
  chrs <- unique(snps[tmp.snps,"chr"])
  
  tmp.fit <- scan1(genoprobs=probs, 
               kinship=Glist[chrs],
               pheno=expr.mrna[,idx], 
               addcovar=addcovar, 
               cores=1, 
               reml=TRUE)
  
  # save LOD scores to results
  result <- rbind(result, tmp.fit$lod)
}
```

