---
title: "Feather Genoprob Features"
author: "Brian S. Yandell"
date: "3/17/2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The motivation is to reduce memory usage. Matt told us that memory is his greatest obstacle, when running QTL browsers because he cannot add more `R/plumber` workers. Also, if you have a cheap Windows laptop you cannot even open many DO `.RData` projects simply because haplotype probability object is too big.
 
The solution would be to store data in some database and load just a slice you need at a time. I experimented with `SQLite` but it was too slow. Then I try feather and it run fine (R script attached – crunching 1000 SNPs at a time). It just does not use `qtl2scan` effectively because the package is not prepared for that.

Task: refactor this Rmd to use DOex and run some examples.

```{r}
suppressPackageStartupMessages({
  library(dplyr)
  library(qtl2geno)
  library(qtl2scan)
  library(qtl2feather)
})
```

```{r}
dirpath <- "~/Documents/Research/rqtl/data/"
if(!file.exists(file <- file.path(dirpath, "DOex.zip"))) {
  dirpath <- paste0("https://raw.githubusercontent.com/rqtl/",
                    "qtl2data/master/DOex")
  file <- file.path(dirpath, "DOex.zip")
}
DOex <- read_cross2(file)
```

Calculate genotype probabilities and convert to allele probabilities

```{r}
pr <- calc_genoprob(DOex, error_prob=0.002)
apr <- genoprob_to_alleleprob(pr)
```

Write feather database of pr.

```{r}
fpr <- feather_genoprob(pr, "pr", "feather")
fapr <- feather_genoprob(apr, "apr", "feather")
```

Methods `names()` and `length()` both work properly.

```{r}
names(fpr)
```

```{r}
length(fpr)
```

Methods `dim()` and `dimnames()` give dimensions across the chromosomes.
Notice that `fpr` has different genotypes for the `X` chromosome.

```{r}
dim(fpr)
```

```{r}
str(dimnames(fpr))
```


### Selecting one chromosome

Selecting a chromosome causes reading from feather database and creation of an array.

```{r}
dim(fapr[["X"]])
```

Alternate form.

```{r}
dim(fapr$X)
```

### Subsetting by ind, chr, mar

Subsetting using `subset(object,ind,chr,mar)` or `[ind,chr,mar]` only adjusts the vector of
individuals, chromosomes and markers, but does not alter the feather database.

```{r}
dim(subset(fapr, ind=1:20, chr=c("2","3")))
```

You can also use `[]` to subset.

```{r}
dim(fapr[1:20, c("2","3")][["3"]])
```

```{r}
f2 <- fapr[,"2"]
f23 <- fapr[,c("2","3")]
fx <- fapr[,"X"]
```

There is a third dimension for markers. However, be careful that if you select a subset of markers that excludes one or more chromosomes, those will be dropped.

```{r}
dim(subset(fapr, mar=1:30))
```

```{r}
dim(fapr[ , , dimnames(fapr)$mar$X[1:20]])
```

### Binding by columns or rows.

Binding by columns or rows will cause creation of a new feather database if input objects arose from different feather databases. However, if they come from the same, then it reuses the one feather database. See `example(cbind.feather_genoprob)` and `example(rbind.feather_genoprob)` with objects having distinct feather databases.

```{r}
newf23 <- cbind(f2,f23)
```

Row bind.

```{r}
f23a <- fapr[1:20, c("2","3")]
f23b <- fapr[40:79, c("2","3")]
f23 <- rbind(f23a, f23b)
```

Subset on markers. This way only extracts the selected `markers` from feather database before creating the array.

```{r}
markers <- dimnames(fapr$X)[[3]][1:10]
dim(fapr[,,markers]$X)
```

This way extracts all markers on `X`, creates the array, then subsets on selected `markers`.

```{r}
markers <- dimnames(fapr$X)[[3]][1:10]
dim(fapr$X[,,markers])
```

### Looking under the hood at `feather_genoprob` object

Here are the names of elements in a `feather_genoprob` object:

```{r}
names(unclass(fapr))
```

```{r}
unclass(fapr)$feather
```

```{r}
sapply(unclass(fapr)[c("ind","chr","mar")], length)
```

A `feather_genoprob` object has all the original information. Thus, it is possible to restore the original object from a `subset` (but not necessarily from a `cbind` or `rbind`). Here is an example.

```{r}
fapr23 <- subset(fapr, chr=c("2","3"))
dim(fapr23)
```

```{r}
dim(feather_genoprob_restore(fapr23))
```

## Time trial simulation

Compare times to create `pr` once and to Read `fpr` 100 times.

```{r}
system.time(pr <- calc_genoprob(DOex, error_prob=0.002))
```

```{r}
system.time({
  for(i in seq(100))
    fpr <- feather_genoprob(pr, "pr", "feather", verbose = FALSE)
})
```

```{r}
tmpfile <- tempfile()
saveRDS(pr, file = tmpfile)
size_pr <- file.size(tmpfile) * 1e-6
unlink(tmpfile)
```

The `pr` object is `r format(object.size(pr), units = "MB")` in `R` or `r size_pr` Mb if saved as RDS file, while the `fpr` object is `r format(object.size(fpr), units = "MB")` in R with additional
`r sum(file.size(unclass(fpr)$feather)) * 1e-6` Mb for saved feather databases.

#### Time to extract chromosome information.

Extract chromosome `"2"` 100 times.

```{r}
system.time({
  for(i in seq(100))
    tmp <- fpr[["2"]]
})
```

Extract first 50 markers from chromosome `"2"` 100 times.
```{r}
markers <- dimnames(fpr)$mar[["2"]][1:50]
system.time({
  for(i in seq(100))
    tmp <- fpr[["2"]]
})
```

